---
title: "Homework 3"
author: "Allison Randy-Cofie"
date: "2022-10-12"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggridges)
library(patchwork)

library(p8105.datasets)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


### Problem 1

#### Read in the data

```{r}
data("instacart")

instacart = 
  instacart %>% 
  as_tibble(instacart)
```

#### Answer questions about the data

This dataset contains `r nrow(instacart)` rows and `r ncol(instacart)` columns, with each row resprenting a single product from an instacart order. Variables include identifiers for user, order, and product; the order in which each product was added to the cart. There are several order-level variables, describing the day and time of the order, and number of days since prior order. Then there are several item-specific variables, describing the product name (e.g. Yogurt, Avocado), department (e.g. dairy and eggs, produce), and aisle (e.g. yogurt, fresh fruits), and whether the item has been ordered by this user in the past. In total, there are `r instacart %>% select(product_id) %>% distinct %>% count` products found in `r instacart %>% select(user_id, order_id) %>% distinct %>% count` orders from `r instacart %>% select(user_id) %>% distinct %>% count` distinct users.

Below is a table summarizing the number of items ordered from aisle. In total, there are 134 aisles, with fresh vegetables and fresh fruits holding the most items ordered by far.

```{r}
instacart %>% 
  count(aisle) %>% 
  arrange(desc(n))
```

Next is a plot that shows the number of items ordered in each aisle. Here, aisles are ordered by ascending number of items.

```{r}
instacart %>% 
  count(aisle) %>% 
  filter(n > 10000) %>% 
  mutate(aisle = fct_reorder(aisle, n)) %>% 
  ggplot(aes(x = aisle, y = n)) + 
  geom_point() + 
  labs(title = "Number of items ordered in each aisle") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

```

Our next table shows the three most popular items in aisles `baking ingredients`, `dog food care`, and `packaged vegetables fruits`, and includes the number of times each item is ordered in your table.

```{r}
instacart %>% 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>%
  group_by(aisle) %>% 
  count(product_name) %>% 
  mutate(rank = min_rank(desc(n))) %>% 
  filter(rank < 4) %>% 
  arrange(desc(n)) %>%
  knitr::kable()
```

Finally is a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week. This table has been formatted in an untidy manner for human readers. Pink Lady Apples are generally purchased slightly earlier in the day than Coffee Ice Cream, with the exception of day 5.

```{r}
instacart %>%
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>%
  group_by(product_name, order_dow) %>%
  summarize(mean_hour = mean(order_hour_of_day)) %>%
  spread(key = order_dow, value = mean_hour) %>%
  knitr::kable(digits = 2)
```



## Problem 2

Load, tidy, and otherwise wrangle the data. Your final dataset should include all originally observed variables and values; have useful variable names; include a weekday vs weekend variable; and encode data with reasonable variable classes. Describe the resulting dataset (e.g. what variables exist, how many observations, etc).
```{r}
accel <- read.csv("./accel_data.csv") %>% 
  janitor::clean_names() %>% 
  pivot_longer(activity_1:activity_1440,
    names_to = "minutes",
    values_to = "activity_count") %>% 
  mutate(minutes = str_sub(minutes, 10),
         minutes = as.integer(minutes),
         day = forcats::fct_relevel(day, "Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday"),
         type_of_day = ifelse(day == "Saturday" | day == "Sunday", "weekend", "weekday"))
```
The dataset has `r nrow(accel)` observations and `r ncol(accel)` variables. The variables in the dataset are minute of the day, activity count, and day of the week.



Traditional analyses of accelerometer data focus on the total activity over the day. Using your tidied dataset, aggregate across minutes to create a total activity variable for each day, and create a table showing these totals. Are any trends apparent?
```{r}
accel %>% 
  group_by(day,week) %>% 
   summarize(tot_activity = sum(activity_count)) %>% 
  pivot_wider(names_from = day,
              values_from = tot_activity) %>% 
  knitr::kable()
```




Accelerometer data allows the inspection activity over the course of the day. Make a single-panel plot that shows the 24-hour activity time courses for each day and use color to indicate day of the week. Describe in words any patterns or conclusions you can make based on this graph.
```{r}
accel %>% 
  ggplot(aes(x=minutes,y=activity_count,group=day_id, color=day))+
  geom_line()+
  labs(title = "24-hour Activity Count for Each Day",
       x = "Time of Day (min)",
       y = "Activity Count")
```


## Problem 3
The goal is to do some exploration of this dataset. To that end, write a short description of the dataset, noting the size and structure of the data, describing some key variables, and indicating the extent to which missing data is an issue. Then, do or answer the following (commenting on the results of each):

Do some data cleaning. Create separate variables for year, month, and day. Ensure observations for temperature, precipitation, and snowfall are given in reasonable units. For snowfall, what are the most commonly observed values? Why?
Make a two-panel plot showing the average max temperature in January and in July in each station across years. Is there any observable / interpretable structure? Any outliers?
Make a two-panel plot showing (i) tmax vs tmin for the full dataset (note that a scatterplot may not be the best option); and (ii) make a plot showing the distribution of snowfall values greater than 0 and less than 100 separately by year.


```{r}
library(p8105.datasets)
data("ny_noaa")
ny_noaa =ny_noaa %>% janitor::clean_names() %>% 
  separate(date, into = c("year","month","day"),sep="-") %>% 
  mutate(
    month = as.numeric(month),
    day = as.numeric(day),
    year = as.numeric(year),
    tmax= as.numeric(tmax),
    tmin= as.numeric(tmin)
  )
          
```

For snowfall the most commonly observed value is 0 since it typically only snows during winter which is only a quarter of the year.



Make a two-panel plot showing the average max temperature in January and in July in each station across years. Is there any observable / interpretable structure? Any outliers?

```{r}
ny_noaa_tmax = 
  ny_noaa %>% 
  group_by( day,month, year) %>% 
  summarize(
    mean_tmax = mean(tmax/10, na.rm=TRUE))
```

```{r}
ny_noaa_tmax %>% 
  filter(month==01 | month==07)%>% 
  ggplot(aes(x = day, y=mean_tmax, color= month))+
  geom_point()+
  facet_grid(. ~ month)
```


Make a two-panel plot showing (i) tmax vs tmin for the full dataset (note that a scatterplot may not be the best option); 
Use density plot. hex plot. for i

```{r}
ny_noaa %>% 
  mutate(tmin=tmin/10,
         tmax=tmax/10) %>% 
  ggplot(aes(x=tmin, y=tmax))+
  geom_point()
```


ii) make a plot showing the distribution of snowfall values greater than 0 and less than 100 separately by year.
```{r}
ny_noaa %>% 
  group_by(id, year) %>% 
  summarize(snow=sum(snow/10)) %>% 
  filter( snow < 100 & snow > 0) %>% 
  ggplot(aes(x=year, y = snow)) +
  geom_boxplot()
```

Use denity or boxplot for this one